---
title: "RecSession"
author: "Kevin Allen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#output: pdf_document
vignette: >
  %\VignetteIndexEntry{relectro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## RecSession

RecSession object read text files (.par, .desen, .desel, etc) and get the information regarding what was done during the recording session. 

We will load the information regarding the recording session that is included within the relectro package as example. 

```{r RecSession}
library(relectro)
## the 3 commands are there to find the data provided with the package on your system.
## you would not have to do this in real life.
clufile<-unlist(strsplit(x=system.file("extdata", 
                                       "jp4298-15022016-0106.clu", 
                                       package = "relectro"), split="/"))
datadir<-paste(clufile[1:length(clufile)-1],sep="/",collapse = "/")
session=strsplit(clufile[length(clufile)],split="\\.")[[1]][1]

## create a RecSession object
## so rs is a RecSession object
rs<-new("RecSession",session=session,path=datadir)

## load configuration data from file
rs<-loadRecSession(rs)
## print information on the recording session
rs
```

RecSession objects are useful because they spare you the trouble of reading and parsing several text files. 

To know which methods are available for 
```{r}
methods(class = "RecSession")
```

As an example, you can get the time intervals when the animal was in a given environment.

```{r}
getIntervalsEnvironment(rs,environment = "lt")
```

