---
title: "relectro"
author: "Kevin Allen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
#output: pdf_document
vignette: >
  %\VignetteIndexEntry{relectro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
relectro is being developped to analyze data from electrophysiological experiments in R. The focus is on data recorded with tetrodes. It is an object-oriented package in which the user create objects to represent recording sessions, spike trains, position tracking, etc. relectro relies on several C functions in computationally intensive segments of the code.

### Install relectro

The most recent source code of relectro on [Github](https://github.com/kevin-allen/relectro)

To download the source code, open a terminal and run :

```git clone https://github.com/kevin-allen/relectro.git```

To install relectro, run the following from the same terminal :

```R CMD build relectro``` 

```R CMD INSTALL relectro```

To use relectro in R:

```{r}
library(relectro)
```

You can list all objects (including functions) in relectro. Here I am printing the first 10.

```{r}
head(ls('package:relectro'),n=10)
```

You can get information regarding these objects or function with `?`

### S4 objects

relectro defines a few S4 objects that represents projects, recording sessions, spike trains, position data, etc. Most functions are actually methods associated with these objects. 

* ElectroProject
* RecSession
* CellGroup
* SpikeTrain
* DatFiles
* Positrack
* HeadDirection
* SpatialProperties1d
* SpatialProperties2d

### Learning by examples

<!-- There is a repository called [relectro_data](https://github.com/kevin-allen/relectro_data). This contains a small data set and some vignette to get you started with using relectro. -->

A good way to learn how to use the code is to look at examples. 
The code used in the paper of Perez-Escobar et al. (2016) is availble for download [here](https://github.com/kevin-allen/prog_perez_escobar_2016).

### Development

If you want to work on the source code, I would recommand reading the book "R packages" by Hadley Wickham. I am using the following packages for development: `devtools`, `roxygen2` and `testthat`.
